### 2015-04-21 R/paper.Rmd
Paste link

---
title:  "The Effect of a Second AIDS Defining Illness on Survival in the MACS"
author: 
- "[Jacob Carey](mailto:jcare15@jhu.edu)"
- "[Matt Murrill](mmurril1@jhmi.edu)"
- "[Kim Kunwha](khkim@jhu.edu)"
date:   "`r Sys.Date()`"
output: pdf_document
abstract: |
  We seek to investigate the effect of a second AIDS defining illness on survival. We included people who are seroprevalent as well as seroconverters. We addressed confounding using propensity scores.
---

```{r chunks, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      message=FALSE)
```

We read in the data.
```{r data}
library(readr)
library(dplyr)
library(ggplot2)
library(lubridate)

# read in horiz data with lots of covariates
covariates <- read_csv("data/macswihs_horiz2013_copy-1.csv") %>%
           # date of death
    mutate(death.date=as.Date("1960-01-01") + FRSTDTHD,
           # date of aids diagnosis
           aids.date=as.Date("1960-01-01") + FRSTAIDD,
           # date of secondary aids diagnosis
           sec.aids.date=as.Date("1960-01-01") + SECAIDD,
           # what was the first aids diagnosis
           frst.aids.diag=AIDSDIAG,
           # what was the second aids diagnosis
           sec.aids.diag=SECAIDSDIAG,
           # did the person have a sec aids diag
           had.sec=year(sec.aids.date) < 2100,
           # did the person die
           died=year(death.date) < 2100,
           age=as.Date("1960-01-01") + DOB,
           exit=ifelse(!died, as.Date("1994-12-31"), death.date),
           censor.date=as.Date("1970-01-01") + as.integer(exit),
           last.alive=as.Date("1960-01-01") + LASTALID,
           censor.date=ifelse(last.alive < censor.date, last.alive, 
                              censor.date),
           dead.at.censor=(death.date < as.Date("1994-12-31"))) %>%
    select(STUDYID, frst.aids.diag, censor.date, sec.aids.diag,
           dead.at.censor)

# read in visit level data
visits <- read_csv("data/hivvert-data04.csv",
                    col_types="iiciiididdiiiiiiiddiiiiiiiiiiiidc") %>%
    # put visits in correct 1,2, 3,... form
    mutate(visit=VISIT / 10,
           visit.date=mdy(VDATE),
           haart.era.visit=year(visit.date) <= 1994) %>%

    # join to covariates
    left_join(covariates, by=c("STUDYID")) %>%

    # keep MACS only
    filter(COHORT==1,
           # keep people with first ADI only
           frst.aids.diag != 0,
           # restrict to haart era visits right now
           haart.era.visit == TRUE) %>%
    mutate(censor.date=as.Date("1970-01-01")+censor.date) %>%
    select(id=STUDYID, center=CENTER, visit:dead.at.censor)
```

```{r fig1}
# therapy <- visits %>%
#     group_by(year(visit.date)) %>%
#     summarise(therapy=list(table(THRPY))) %>%
#     unnest(therapy)

theme_set(theme_classic())

visits %>%
    group_by(id) %>%
    summarise(sec.aids=min(sec.aids.diag)) %>%
    mutate(sec.aids=plyr::revalue(as.character(sec.aids),
                                  c("0"="No Aids",
                                    "1"="PCP",
                                    "2"="KS",
                                    "3"="NHL",
                                    "4"="Neuro",
                                    "5"="Bacteria OI",
                                    "6"="Fungal OI",
                                    "7"="Protozoa",
                                    "8"="CMV/HSV",
                                    "9"="Wasting",
                                    "10"="Recurrent PNA",
                                    "11"="Cervical Ca")),
           warn_missing=FALSE) %>%
ggplot(., aes(x=sec.aids)) + geom_histogram() +
    xlab("Type of Secondary Aids Diagnosis") +
    ylab("Count") +
    ggtitle("Figure 1") +
    theme(axis.text.x = element_text(angle=-90)) +
    scale_x_discrete(limits=k
```

```{r naive-survival}
# http://cran.r-project.org/web/packages/survival/vignettes/timedep.pdf
```