### 2015-05-05 paper.Rmd
First draft of practicum done

---
title:  "The Effect of a Second AIDS Defining Illness on Survival in the MACS"
author: 
- "[Jacob Carey](mailto:jcare15@jhu.edu)"
- "[Matt Murrill](mmurril1@jhmi.edu)"
- "[Kim Kunwha](khkim@jhu.edu)"
date:   "`r Sys.Date()`"
output: pdf_document
abstract: |
  We seek to investigate the effect of a second AIDS defining illness on survival. We included people who are seroprevalent as well as seroconverters. We addressed confounding using propensity scores.
---

```{r chunks, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      message=FALSE,
                      results='asis')
```

```{r libs}
library(readr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(survival)
theme_set(theme_classic())
```

We read in the data.
```{r data}
# read in horiz data with lots of covariates
covariates <- read_csv("data/macswihs_horiz2013_copy-1.csv", 
                       progress=FALSE) %>%
           # date of death
    mutate(death.date=as.Date("1960-01-01") + FRSTDTHD,
           # date of aids diagnosis
           aids.date=as.Date("1960-01-01") + FRSTAIDD,
           # date of secondary aids diagnosis
           sec.aids.date=as.Date("1960-01-01") + SECAIDD,
           # what was the first aids diagnosis
           frst.aids.diag=AIDSDIAG,
           # what was the second aids diagnosis
           sec.aids.diag=SECAIDSDIAG,
           # did the person have a sec aids diag
           had.sec=year(sec.aids.date) < 2100,
           # did the person die
           died=year(death.date) < 2100,
           age=as.Date("1960-01-01") + DOB,
           exit=ifelse(!died, as.Date("1994-12-31"), death.date),
           censor.date=as.Date("1970-01-01") + as.integer(exit),
           last.alive=as.Date("1960-01-01") + LASTALID,
           censor.date=ifelse(last.alive < censor.date, last.alive, 
                              censor.date),
           censor.date=as.Date("1970-01-01")+censor.date,
           dead.at.censor=(death.date < as.Date("1994-12-31"))) %>%
    select(STUDYID, 
           aids.date,
           sec.aids.date,
           frst.aids.diag, 
           censor.date, 
           sec.aids.diag,
           dead.at.censor, 
           death.date) %>%
    filter(year(aids.date) < 2100)

# read in visit level data
visits <- read_csv("data/hivvert-data04.csv", progress=FALSE,
                    col_types="iiciiididdiiiiiiiddiiiiiiiiiiiidc") %>%
    # keep MACS only
    filter(COHORT==1) %>%
    select(STUDYID) %>%
    distinct(STUDYID)

covariates <- covariates %>%
    inner_join(visits)
```

```{r fig1}
visits <- read_csv("data/hivvert-data04.csv", progress=FALSE,
                    col_types="iiciiididdiiiiiiiddiiiiiiiiiiiidc") %>%
    # put visits in correct 1,2, 3,... form
    mutate(visit=VISIT / 10,
           visit.date=mdy(VDATE),
           haart.era.visit=year(visit.date) <= 1994) %>%

    # join to covariates
    left_join(covariates, by=c("STUDYID")) %>%

    # keep MACS only
    filter(COHORT==1,
           # keep people with first ADI only
           frst.aids.diag != 0,
           # restrict to haart era visits right now
           haart.era.visit == TRUE) %>%
    select(id=STUDYID, center=CENTER, visit:dead.at.censor)

# only for pre haart!
# need to do again for all data
data.plot <- visits %>%
    group_by(id) %>%
    summarise(sec.aids=min(sec.aids.diag)) %>%
    mutate(sec.aids=plyr::revalue(as.character(sec.aids),
                                  c("0"="No Aids",
                                    "1"="PCP",
                                    "2"="KS",
                                    "3"="NHL",
                                    "4"="Neuro",
                                    "5"="Bacteria OI",
                                    "6"="Fungal OI",
                                    "7"="Protozoa",
                                    "8"="CMV/HSV",
                                    "9"="Wasting",
                                    "10"="Recurrent PNA",
                                    "11"="Cervical Ca"),
                                  warn_missing=FALSE))
ggplot(data.plot, aes(x=sec.aids)) + 
    geom_histogram() +
    xlab("Type of Secondary Aids Diagnosis") +
    ylab("Count") +
    ggtitle("Figure 1") +
    scale_x_discrete(limits=names(sort(-table(data.plot$sec.aids)))) +
    theme(axis.text.x = element_text(angle=-90))
```

We fit a Cox proportional hazards model to the pre haart era data. We allowed secondary aids diagnosis to be time varying. 
```{r naive-survival}
# http://cran.r-project.org/web/packages/survival/vignettes/timedep.pdf
# https://stat.ethz.ch/R-manual/R-devel/library/survival/html/coxph.html
data <- covariates %>%
    filter(year(aids.date) <= 1994) %>%
    rowwise() %>%
    transmute(start=aids.date,
              stop=min(censor.date, as.Date("1994-12-31")),
              sec.aids=sec.aids.date,
              status=(death.date <= as.Date("1994-12-31"))) %>%
    mutate_each(funs(as.numeric), start, stop, sec.aids) %>%
    # there are a few obs coming up where secondary and primary are 
    # diagnosed at the same time.
    # let's remove those.
    filter(stop > start,
           !is.na(sec.aids))

pre.sec <- data %>%
    rowwise() %>%
    mutate(stop=min(stop, sec.aids),
           sec.aids=0) %>%
    # remove those with same date diag of first and sec aids
    filter(start < stop)

post.sec <- data %>%
    filter(sec.aids < stop) %>%
    mutate(start=sec.aids,
           sec.aids=1)

time.varying <- bind_rows(pre.sec, post.sec)

naive.model <- coxph(Surv(start, stop, status) ~ sec.aids,
                     data=time.varying)

knitr::kable(coef(summary(naive.model)),
             caption="Naive Model (All Data)")
```

```{r all-data}
# read in horiz data with lots of covariates
covariates <- read_csv("data/macswihs_horiz2013_copy-1.csv") %>%
           # date of death
    mutate(death.date=as.Date("1960-01-01") + FRSTDTHD,
           # date of aids diagnosis
           aids.date=as.Date("1960-01-01") + FRSTAIDD,
           # date of secondary aids diagnosis
           sec.aids.date=as.Date("1960-01-01") + SECAIDD,
           # what was the first aids diagnosis
           frst.aids.diag=AIDSDIAG,
           # what was the second aids diagnosis
           sec.aids.diag=SECAIDSDIAG,
           # did the person have a sec aids diag
           had.sec=year(sec.aids.date) < 2100,
           # did the person die
           died=year(death.date) < 2100,
           dob=as.Date("1960-01-01") + DOB,
           age.at.aids=difftime(aids.date, dob, unit="weeks") / 52.25,
           age.at.aids=as.numeric(age.at.aids),
           exit=death.date,
           censor.date=as.Date("1970-01-01") + as.integer(exit),
           last.alive=as.Date("1960-01-01") + LASTALID,
           censor.date=ifelse(last.alive < censor.date, last.alive, 
                              censor.date),
           censor.date=as.Date("1970-01-01")+censor.date) %>%
    select(STUDYID, 
           aids.date,
           sec.aids.date,
           frst.aids.diag, 
           censor.date, 
           sec.aids.diag,
           death.date,
           race=RACE,
           age.at.aids) %>%
    filter(year(aids.date) < 2100) %>%
    mutate(haart.era=(year(aids.date) > 1993),
           aids.year=year(aids.date),
           aids.month=month(aids.date),
           race=ifelse(race == 1, "white", "non-white"))

# what are the important confoudners that we can include for the PH model?
# what about propensity score?

# what about therapy?
visits <- read_csv("data/hivvert-data04.csv", progress=FALSE,
                    col_types="iiciiididdiiiiiiiddiiiiiiiiiiiidc") %>%
    # keep MACS only
    filter(COHORT == 1) %>% 
    group_by(STUDYID) %>%
    summarise(drug.use=sum(POTUSE, CRACKUSE, COCUSE, HERUSE, AMPHUSE, 
                           POPUSE, na.rm=TRUE) > 1,
              depression=mean(CESD, na.rm=TRUE) < ifelse(drug.use, 
                                                         16, 23)) %>%
    ungroup() %>%
    # remove the 8 missing depression status
    filter(!is.na(depression))

# cd4 closest to primary AIDS diagnosis
visits <- read_csv("data/hivvert-data04.csv", progress=FALSE,
                    col_types="iiciiididdiiiiiiiddiiiiiiiiiiiidc") %>%
    filter(COHORT == 1, !is.na(CD4N)) %>%
    select(STUDYID, cd4=CD4N, VDATE) %>%
    mutate(VDATE=mdy(VDATE),
           VDATE=as.Date(VDATE)) %>%
    inner_join(covariates, by=c("STUDYID"="STUDYID")) %>%
    select(STUDYID, cd4, VDATE, aids.date) %>%
    mutate(visit.aids.diff=difftime(VDATE, aids.date)) %>%
    group_by(STUDYID) %>%
    filter(abs(visit.aids.diff) == min(abs(visit.aids.diff))) %>%
    select(STUDYID, cd4) %>%
    ungroup() %>%
    right_join(visits)

# add data together
covariates <- covariates %>%
    inner_join(visits) 
```

We also created a naive model that includes **all** MACS data,
but does not adjust for any confounders, or the pre / post HAART era.

```{r all-naive-model}
data <- covariates %>%
    rowwise() %>%
    mutate(start=aids.date,
           stop=censor.date,
           sec.aids=sec.aids.date,
           status=(death.date < as.Date("2100-01-01"))) %>%
    mutate_each(funs(as.numeric), start, stop, sec.aids) %>%
    # there are a few obs coming up where secondary and primary are 
    # diagnosed at the same time.
    # let's remove those.
    filter(stop > start,
           !is.na(sec.aids))

pre.sec <- data %>%
    rowwise() %>%
    mutate(stop=min(stop, sec.aids),
           sec.aids=0) %>%
    # remove those with same date diag of first and sec aids
    filter(start < stop)

post.sec <- data %>%
    filter(sec.aids < stop) %>%
    mutate(start=sec.aids,
           sec.aids=1)

time.varying <- bind_rows(pre.sec, post.sec)

naive.model <- coxph(Surv(start, stop, status) ~ sec.aids,
                     data=time.varying)

knitr::kable(coef(summary(naive.model)),
             caption="Naive Model")
```

Next, we adjusted for several confounders, such as Race, Age at baseline, CD4 at AIDS diagnosis, year of AIDS diagnosis, and drug use.

```{r all-adjusted}
adjusted.model <- coxph(Surv(start, stop, status) ~ sec.aids +
                        factor(race) + age.at.aids + drug.use +
                        aids.year + depression + cd4,
                        data=time.varying)

adjusted.model.ph <- cox.zph(adjusted.model)

knitr::kable(coef(summary(adjusted.model)),
             caption="PH Model Adjusted for Confounders")

knitr::kable(cox.zph(adjusted.model)$table,
             caption="PH Assumption Test")
```

Finally, we use propensity score matching to account for confounding bias.

```{r propensity-score}
caliper <- 0.07
set.seed(90210)
match.set <- MatchIt::matchit(sec.aids ~ factor(race) + 
                              age.at.aids + drug.use +
                              aids.year + depression + cd4,
                              data=time.varying,
                              distance="logit",
                              method="nearest",
                              caliper=caliper,
                              discard="both")

# ## let's check out the diagnostics

# ## numerical check
match.check <- summary(match.set, standardize=TRUE)

# ## graphical check
plot(match.set, type="hist")
# plot(match.set, type="jitter", interactive=FALSE)

matched.dat <- MatchIt::match.data(match.set,"all")

ph.match <- coxph(Surv(start, stop, status) ~ sec.aids +
                  factor(race) + age.at.aids + drug.use +
                  aids.year + depression + cd4,
   				  data=matched.dat,
                  method="efron")
```

There were `r length(unique(matched.dat$STUDYID))` individuals in the 
matched data set of the orginal `r nrow(data)`. There are 
`r sum(matched.dat$sec.aids)` events in the matched data set.
We perform a check on the data and see the following summaries. We used a
caliper of `r caliper`.

```{r}
knitr::kable(match.check$sum.all[,1:4],
             caption="Summary of balance for all data")
knitr::kable(match.check$sum.matched[,1:4],
             caption="Summary of balance for matched data")
knitr::kable(match.check$reduction,
             caption="Percent Balance Improvement")
knitr::kable(match.check$nn,
             caption="Sample Sizes")
```

In recalculating our proportional hazards model using propensity score
in order to handle confounding bias, we see the following results.

```{r}
knitr::kable(coef(summary(ph.match)),
             caption="PH Model (Propensity Scores)")

knitr::kable(cox.zph(ph.match)$table,
             caption="PH Assumption Test")
caliper <- 0.5
```

Next, we performed the same PH model using propensity scores, but with a higher
caliper (`r caliper`).

```{r propensity-score-2}
set.seed(90210)
match.set <- MatchIt::matchit(sec.aids ~ factor(race) + 
                              age.at.aids + drug.use +
                              aids.year + depression + cd4,
                              data=time.varying,
                              distance="logit",
                              method="nearest",
                              caliper=caliper,
                              discard="both")

# ## let's check out the diagnostics

# ## numerical check
match.check <- summary(match.set, standardize=TRUE)

# ## graphical check
plot(match.set, type="hist")
# plot(match.set, type="jitter", interactive=FALSE)

matched.dat <- MatchIt::match.data(match.set,"all")

ph.match <- coxph(Surv(start, stop, status) ~ sec.aids +
                  factor(race) + age.at.aids + drug.use +
                  aids.year + depression + cd4,
   				  data=matched.dat,
                  method="efron")
```

There were `r length(unique(matched.dat$STUDYID))` individuals in the 
new matched data set of the orginal `r nrow(data)`. There are 
`r sum(matched.dat$sec.aids)` events in the matched data set.
We perform a check on the data and see the following summaries.

```{r}
knitr::kable(match.check$sum.all[,1:4],
             caption="Summary of balance for all data")
knitr::kable(match.check$sum.matched[,1:4],
             caption="Summary of balance for matched data")
knitr::kable(match.check$reduction,
             caption="Percent Balance Improvement")
knitr::kable(match.check$nn,
             caption="Sample Sizes")
```

In our second PH model with propensity scores, we see the following results.

```{r}
knitr::kable(coef(summary(ph.match)),
             caption="PH Model (Propensity Scores)")

knitr::kable(cox.zph(ph.match)$table,
             caption="PH Assumption Test")
caliper <- 0.5
```

Finally, we used a subclassification method for our propensity score.

```{r subclass}
sub.set <- MatchIt::matchit(sec.aids ~ factor(race) + 
                            age.at.aids + drug.use +
                            aids.year + depression + cd4,
                            data=time.varying,
                            method="subclass",
                            sub.by="all",
                            subclass=5)

# summary(sub.set,standardize=T)
plot(sub.set,type="jitter",interactive=F)
# plot(summary(sub.set,standardize=T),interactive=F)

sub.dat <- MatchIt::match.data(sub.set, "all")
ph.match <- coxph(Surv(start, stop, status) ~ sec.aids + strata(subclass),
   				  data=sub.dat,
                  method="efron")
knitr::kable(coef(summary(ph.match)),
             caption="PH Model (Propensity Scores)")

knitr::kable(cox.zph(ph.match)$table,
             caption="PH Assumption Test")
```

```{r graphs}
plot(survfit(ph.match), main="Final PH Model Using Subclassification")
```